---
title: "CIS 428 Text Mining D3"
author: "Team 3"
date: "2026-04-21"
output: pdf_document
---
Three Text Mining Methods for classification, & one for topic modeling are used here.
First classification using Naive Bayes.
```{r}

#load required packages
library(tidyverse)
library(tm)
library(e1071)
library(gmodels)
library(wordcloud)

# 1. LOAD DATASET.
# Load your cleaned dataset
data <- read.csv("BA_Reviews_Cleaned.csv", stringsAsFactors = FALSE)
# Check structure
str(data)
summary(data)

# Select only required columns (adjust column names if needed)
data <- data %>%
  select(Review, Sentiment) %>%
  na.omit()

View(data)
# Rename columns to better names.
names(data) <- c("text", "score")

# Convert score to factor
data$score <- factor(data$score)

# Check distribution
table(data$score)
```
The clean dataset was loaded and only the two required columns were selected: Review & Sentiment. Column names were changed and factored.

```{r}
# 2. CREATE CORPUS
corpus <- VCorpus(VectorSource(data$text))

# Create Document-Term Matrix (DTM)
dtm <- DocumentTermMatrix(corpus, control = list(
  tolower = TRUE,
  removeNumbers = TRUE,
  stopwords = TRUE,
  removePunctuation = TRUE,
  stemming = TRUE
))
inspect(dtm)

# Splitting Data into Train & Test 80/20.

m <- nrow(data)

train_data <- data[1:round(.8 * m), ]
test_data  <- data[(round(.8 * m) + 1):m, ]

train_labels <- train_data$score
test_labels  <- test_data$score

# Create WordCloud.
positive <- subset(train_data, score == "Positive")
negative <- subset(train_data, score == "Negative")

wordcloud(positive$text, max.words = 40, scale = c(3, 0.5))
wordcloud(negative$text, max.words = 40, scale = c(3, 0.5))

```
A Corpus was created.Then a dtm. After that train and test data were divided into 80:20- labels were also trained
similarly for the original data.. Wordcloud was made based on score being "positive" or "negative."

```{r}
# 3. SPLIT DTM. Train & Test Naive Bayes Model

n <- nrow(dtm)

dtm_train <- dtm[1:round(.8 * n), ]
dtm_test  <- dtm[(round(.8 * n) + 1):n, ]

# Convert Counts to Binary
convert_counts <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
  }
train_binary <- apply(dtm_train, MARGIN = 2, convert_counts)
test_binary  <- apply(dtm_test, MARGIN = 2, convert_counts)

# Train Naive Bayes Model
nb_model <- naiveBayes(as.matrix(train_binary), train_labels)

# Prediction
nb_pred <- predict(nb_model, as.matrix(test_binary))

```
DTM was split and counts converted to binary. Then a model was made using naive bayes and used for predicting.

```{r}
# Evaluation with Confusion Matrix
CrossTable(nb_pred, test_labels,
           prop.chisq = FALSE,
           prop.t = FALSE,
           prop.r = FALSE,
           dnn = c("Predicted", "Actual"))

#Laplace Smoothing
nb_model_laplace <- naiveBayes(as.matrix(train_binary), train_labels, laplace = 1)

nb_pred_laplace <- predict(nb_model_laplace, as.matrix(test_binary))

CrossTable(nb_pred_laplace, test_labels,
           prop.chisq = FALSE,
           prop.t = FALSE,
           prop.r = FALSE,
           dnn = c("Predicted", "Actual"))

```
Evaluation was done with crosstable. The result of confusion matrix shows that:
Of the 247 actual Negative reviews, 135 (54.7%) were correctly predicted as Negative
Of the 111 actual Neutral reviews, 31 (27.9%) were incorrectly predicted as Negative
Of the 382 actual Positive reviews, 14 (3.7%) were incorrectly predicted as Negative.
Laplace smoothing didn't help with probability, predicting 739/740 as neutral.

Now, using second classification algorithm, KNN.
```{r}
# 1. Data loading process.
# Load required libraries
library(tm)        # Text preprocessing
library(class)     # KNN model
library(caret)     # Data splitting and evaluation
library(dplyr)     # Data manipulation

# 1. LOAD DATASET 
# Load cleaned dataset
data <- read.csv("BA_Reviews_Cleaned.csv", stringsAsFactors = FALSE)

# Select only required columns (adjust column names)
data <- data %>%
  select(Review, Sentiment) %>%
  na.omit()

# Rename columns to better names
names(data) <- c("text", "score")

# Convert score to factor
data$score <- factor(data$score)

# Check distribution
table(data$score)

```



```{r}
# 2. TEXT PREPROCESSING FOR KNN ()
# Convert text messages to corpus
corpus <- VCorpus(VectorSource(data$text))

# Apply text transformations
corpus <- tm_map(corpus, content_transformer(tolower))      # Convert to lowercase
corpus <- tm_map(corpus, removePunctuation)                  # Remove punctuation
corpus <- tm_map(corpus, removeNumbers)                      # Remove numbers
corpus <- tm_map(corpus, removeWords, stopwords("english"))  # Remove stopwords
corpus <- tm_map(corpus, stripWhitespace)                    # Remove extra spaces

# Remove custom words (add any domain-specific words to remove) (opt.)
custom_stopwords <- c("british", "airways", "flight", "ba")
corpus <- tm_map(corpus, removeWords, custom_stopwords)

# Convert corpus to Document-Term Matrix (DTM)
dtm <- DocumentTermMatrix(corpus)

# Remove sparse terms (adjust threshold as needed)

dtm <- removeSparseTerms(dtm, 0.99)  # Keeps terms appearing in at least 1% of docs
# 

# Convert DTM to a data frame
data_matrix <- as.data.frame(as.matrix(dtm))

# Add labels back to the dataset
data_matrix$score <- data$score

# Check the dimensions of the feature matrix
cat("Feature matrix dimensions:", dim(data_matrix), "\n")
cat("Number of features (words):", ncol(data_matrix) - 1, "\n")
# View(data_matrix)
```
A corpus was created. Then text transformations were performed on the data which was obtained pre-processed earlier. 
Custom stop words not necessary for sentiment analysis were removed. After that, it was converted to dtm and, terms appearing in less than 1% of documents were removed to reduce dimensionality.

```{r}
#3. SPLIT DATA INTO TRAINING AND TESTING SETS
set.seed(123)  # For reproducibility
trainIndex <- createDataPartition(data_matrix$score, p = 0.8, list = FALSE)
trainData <- data_matrix[trainIndex, ]
testData <- data_matrix[-trainIndex, ]

# Extract features and labels
train_text <- trainData[, -ncol(trainData)]  # All columns except last (score)
train_labels <- trainData$score

test_text <- testData[, -ncol(testData)]
test_labels <- testData$score

#scale data.
train_text <- scale(train_text)
test_text  <- scale(test_text)

```
Data was split into training and testing sets 80/20. Features and labels were extracted and scaled.

```{r}
# 4. TRAIN AND EVALUATE KNN WITH DIFFERENT K VALUES

# Method 1: Simple KNN with k=5
knn_pred_k5 <- knn(train = train_text, 
                   test = test_text, 
                   cl = train_labels, 
                   k = 5)

# Evaluate k=5 model
conf_matrix_k5 <- table(Predicted = knn_pred_k5, Actual = test_labels)
print("Confusion Matrix for k=5:")
print(conf_matrix_k5)

accuracy_k5 <- sum(diag(conf_matrix_k5)) / sum(conf_matrix_k5)
cat("Model Accuracy (k=5):", round(accuracy_k5 * 100, 2), "%\n")

```
Using k = 5 model, accuracy was found to be 59.62% in this model.


```{r}
# 4.
# Method 2: Cross-validation to find optimal k
# Create trainControl object for cross-validation
train_control <- trainControl(method = "cv",       # Cross-validation method
                              number = 10,         # 10-fold cross-validation
                              search = "grid",     # Search through the specified grid
                              verboseIter = FALSE) # Set to TRUE to see progress

# Set the grid for hyperparameter tuning (try different k values)
tune_grid <- expand.grid(k = c(3, 5, 7, 9, 11, 15, 21, 31))

# KNN can be slow with large feature sets
# So, scaling features for better performance
knn_model <- train(x = train_text,                    # Features for training
                   y = train_labels,                  # Labels for training
                   method = "knn",                    # KNN algorithm
                   trControl = train_control,         # Cross-validation control
                   tuneGrid = tune_grid,              # Grid search for 'k'
                   preProcess = c("center", "scale"), # Preprocessing: centering and scaling
                   metric = "Accuracy")               # Metric for model evaluation

# Print the trained model details
print(knn_model)

# Plot accuracy vs k value
plot(knn_model, main = "KNN Accuracy by k Value")

# Get the best k value
best_k <- knn_model$bestTune$k
cat("\nBest k value from cross-validation:", best_k, "\n")
cat("Best accuracy:", round(max(knn_model$results$Accuracy) * 100, 2), "%\n")

# Make predictions with the best tuned model
knn_pred_best <- predict(knn_model, newdata = test_text)

# Evaluate best model
conf_matrix_best <- table(Predicted = knn_pred_best, Actual = test_labels)
print("Confusion Matrix for Best Model:")
print(conf_matrix_best)

accuracy_best <- sum(diag(conf_matrix_best)) / sum(conf_matrix_best)
cat("Best Model Accuracy:", round(accuracy_best * 100, 2), "%\n")

```

Cross-Validation method was used to find optimal k value. A plot was made that showed KNN accuracy by k value. Best K value was
found to be 3, with accuracy of 56.16%. Then, the best tuned model was evaluated. After that the best model accuracy was determined.

The third classification algorithm in use is Lexicon Based Sentiment Analysis. 

```{r}
# LEXICON BASED SENTIMENT ANALYSIS

# Data and Packages loading step.
# Use "BA_Reviews_Cleaned.csv" dataset.

library(tidytext)
library(textdata)
library(dplyr)
library(ggplot2)
library(tidyr)

# Load cleaned dataset 
data <- read.csv("BA_Reviews_Cleaned.csv", stringsAsFactors = FALSE)

# Select columns
data <- data %>%
  select(Review, Sentiment) %>%
  na.omit()

# Rename
names(data) <- c("text", "score")

# Add ID for each review
data <- data %>%
  mutate(id = row_number())

# Tokenize
tidy_reviews <- data %>%
  unnest_tokens(word, text)

#remove stopwords.
tidy_reviews <- tidy_reviews %>%
  anti_join(stop_words, by = "word")
View(tidy_reviews)
str(tidy_reviews)

custom_stopwords <- data.frame(
  word = c("flight", "british", "airways", "ba", "airline",
           "london", "heathrow", "just", "also", "really",
           "one", "get", "got", "time", "will")
)
tidy_reviews <- tidy_reviews %>%
  anti_join(custom_stopwords, by = "word")

```
The dataset was tokenized cleaned by removing stop words. 



```{r}

sentiments
get_sentiments("afinn")
get_sentiments("bing")
get_sentiments("nrc")

# BING SENTIMENT 
bing_sentiment <- tidy_reviews %>%
  inner_join(get_sentiments("bing"), by = "word") %>%
  count(id, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

# AFINN SENTIMENT
afinn_sentiment <- tidy_reviews %>%
  inner_join(get_sentiments("afinn"), by = "word") %>%
  group_by(id) %>%
  summarise(sentiment = sum(value), .groups = "drop") %>%
  mutate(method = "AFINN")

# NRC SENTIMENT
nrc_sentiment <- tidy_reviews %>%
  inner_join(get_sentiments("nrc"), by = "word") %>%
  filter(sentiment %in% c("positive", "negative")) %>%
  count(id, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative,
         method = "NRC")

```
Three sentiments were applied to the tidy token data.


```{r}
# COMBINE ALL METHODS
combined_sentiment <- bind_rows(
  afinn_sentiment,
  bing_plot,
  nrc_sentiment
)

# Plot comparison
ggplot(combined_sentiment, aes(id, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y") +
  labs(title = "Sentiment Comparison Across Methods")

# WORD CLOUD 

library(wordcloud)

tidy_reviews %>%
  anti_join(stop_words) %>%
  count(word, sort = TRUE) %>%
  with(wordcloud(word, n, max.words = 100))



```

The results of all three methods was plotted in a side by side bar plot.
Then, word cloud showed most common words excluding stop words.

2. Text mining method- Topic Modeling using LDA 

```{r}
# Data preprocessing step.
# Load required libraries
library(tm)
library(topicmodels)
library(tidytext)
library(dplyr)
library(ggplot2)

# 1. LOAD DATASET
data <- read.csv("BA_Reviews_Cleaned.csv", stringsAsFactors = FALSE)

# Select required columns
data <- data %>%
  select(Review, Sentiment) %>%
  na.omit()

# Rename columns
names(data) <- c("text", "score")

# 2. CREATE CORPUS & DTM
corpus <- VCorpus(VectorSource(data$text))

# Text preprocessing
corpus <- tm_map(corpus, content_transformer(tolower))
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, removeNumbers)
corpus <- tm_map(corpus, removeWords, stopwords("english"))
corpus <- tm_map(corpus, stripWhitespace)

# Create DTM
dtm <- DocumentTermMatrix(corpus)

# Remove sparse terms
dtm <- removeSparseTerms(dtm, 0.99)

# Remove empty rows
rowTotals <- apply(dtm, 1, sum)
dtm <- dtm[rowTotals > 0, ]

```

```{r}
# 3. BUILD LDA MODEL
# Set number of topics (adjust if needed)
k <- 3
lda_model <- LDA(dtm, k = k, control = list(seed = 123))

# 4. VIEW TOP TERMS PER TOPIC
topics <- terms(lda_model, 10)  # top 10 words per topic
print(topics)

```
Topic 1 labelled as : Overall, flight experience and service quality.
Topic 2 labelled as : Airline Operations & Business Class Experience.
Topic 3 labelled as : Airport Experience, Seating & Travel Time.

Some common terms such as ‘flight’ appeared across multiple topics, indicating their general importance across all reviews rather than topic-specific relevance.


```{r}
# 5. TIDY FORMAT FOR VISUALIZATION
tidy_lda <- tidy(lda_model, matrix = "beta")

top_terms <- tidy_lda %>%
  group_by(topic) %>%
  slice_max(beta, n = 10) %>%
  ungroup() %>%
  arrange(topic, -beta)

# Plot top words per topic
ggplot(top_terms, aes(x = reorder_within(term, beta, topic), 
                      y = beta, 
                      fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~topic, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  labs(title = "Top Terms per Topic",
       x = "Terms",
       y = "Probability")

# 6. DOCUMENT-TOPIC DISTRIBUTION

doc_topics <- tidy(lda_model, matrix = "gamma")

head(doc_topics, 25)

```
Output from gamma implies that each document was associated with multiple topics with varying probabilities. It indicates
that reviews often discuss multiple aspects of the flight experience.


```{r}


```











